---
title: Syllabus
format: 
  html: default
  pdf: default
date-modified: 2023-08-26
email-obfuscation: javascript

---

## Contact Information

Instructor: Vitaly Druker  <br>
Email:  [vdruker@ccny.cuny.edu](mailto:vdruker@ccny.cuny.edu) <br> 
Office Hours:  TBD

## Text Book

Mendenhall, W., Beaver, R. J., & Beaver, B. M. (2019). Introduction to probability and statistics (15th ed.). CENGAGE Learning Custom Publishing.

[Available at CCNY Store](https://ccny.textbookx.com/institutional/?action=browse#books/3838149)

ISBN: 9781337554428

## Topics

Descriptive statistics and frequency histograms; measures of location and dispersion; elementary probability; permutations and combinations; multiplication rule and conditional probability; Bayes’ Theorem; independent events; random variables, expected values; applications to binomial, uniform and normal distributions; the Central Limit Theorem; linear regression and least squares.

## Grades

- Quizzes (6/~10) 20%
- Exams (2) 20% Each
- Final Exam 40%
- Extra credit Project (5%)

## Overview

This course is designed to give an overview of the key concepts in probability and statistics. We will cover new material every class according the [tentative schedule](#schedule) below. There will be a short quiz once a week (generally Tuesdays) unless otherwise stated (we will have about 10 during the semester). Only the top 6 will count towards your grade - makeups will not be available.
Suggested homework problem will be assigned.

We will primarily use calculators in this course for arithmatic. However, a number of calculations are not possible to do on a calculator or take too long (adding 100 numbers for example). We will therefore learn how to do a lot of the work in the Excel as well. If there is interest/time I will also show how some work may be done using R (a programming language).

## Calculators

You are required to obtain and know how to use a scientific calculator capable of handling
square-root expressions and exponential expressions (e.g. √2, 1.063, etc.). An affordable calculator that I recommend is the TI-30X IIS. It sells for $15.99 on Amazon. Graphing calculators can also be borrowed from the library for free. Your cell phone cannot be used as a calculator during a quiz/exam.

## Attendance

Every lecture will be held in person unless announced otherwise. Please be sure to attend every lecture so you are better prepared for the homework/quizes. If any unexpected event occurs, be prepared to meet remotely via Blackboard Collaborate Ultra. Such a switch will be announced in advance.

## Academic Integrity

From [The City College’s website](https://www.ccny.cuny.edu/about/integrity):

> Academic integrity is an essential part of the pursuit of truth, and of your education. We are all are all responsible for maintaining academic integrity at City College – it is the rock on which the value of your degree is built. <br> 
If you cheat on a test or plagiarize by using someone else’s work or ideas, you defeat the purpose of your education. In addition, academic dishonesty is prohibited in the City University of New York, and is punishable by failing grades, suspension and expulsion.

## Disabilities

Under the Americans with Disabilities Act, all members of the campus community are entitled to equal access to the programs and activities of The City College of New York. If you have (or think that you might have) a disability that may impact your participation in the activities, coursework, or assessment of this course, you may be entitled to accommodations through the Accessability Center/Student Disability Services. You can contact them at 212-650-5913, or at [disabilityservices@ccny.cuny.edu](mailto:disabilityservices@ccny.cuny.edu).

## Schedule

```{r}
#| include: false
library(readxl)
library(here)
library(gt)
```

```{r}
#| echo: false
schedule_raw <- read_xlsx(here('resources','semester-planning.xlsx'))

schedule_raw |> 
    subset(select = c(
        `Lecture Number`, Date, 
        `Lecture Topic`)) |> 
        gt() |> 
    tab_style(
    style = cell_fill(color = "lightgrey"),
    locations = cells_body(
      rows = is.na(`Lecture Topic`)
    )
  ) |> 
  fmt_date(columns = `Date`,
  date_style = "MEd") |>  
  sub_missing(missing_text = "")
```