---
title: "Homework Chapter 6"
format: html
---

```{r}
#| include: false
g <- glue::glue
library(ggplot2)
sapply(list.files(here::here('R'), full.names = T), source)
```

## Section 1

### Q18

```{r}
curve(dunif(x, min = 25, max = 50), from = 20, to = 55, col = "blue", lwd = 3, ylab = "f(x)", xaxt = 'n')
axis(side = 1, at = seq(20, 55, by = 5))
```

$$
f(x) = \frac{1}{b-a} = \frac{1}{50 - 25} = \frac{1}{25}
$$

$$
P(x > 45) = (50 - 45) * \frac{1}{25} = 5/25 = .2
$$

$$
P(35 < x < 45) = (45 - 35) * \frac{1}{25} = 10/25 = .4
$$

$$
P(x < 40) = (40 - 25) * \frac{1}{25} = 15/25 = .6
$$

### Q19


```{r}
curve(dexp(x, rate = 1/365), from = 0, to = 365*2, col = "blue", lwd = 3, ylab = "f(x)", xaxt = 'n')
axis(side = 1, at = c(0, 180, 365, 356*2))
```

$$
P(x < 180) = 1 - e^{-\frac{1}{365} \times 180} = .39
$$

$$
P(x > 365) = e^{-365/365} = .37
$$


- P(A) = battery 1 lasts
- P(B) = battery 2 lasts

Assume A and B are independent

$$
P(A \cap B) = P(A)P(B) = .37 * .37 = 0.14
$$

## Section 2

### Q46-48

46

```{r}
pnorm(13, 12.9, 1.6) - pnorm(12, 12.9, 1.6)
```


47
```{r}
1 - pnorm(16, 14.1, 1.6)
```

48


The chance of seeing that in a boy is:
```{r}
pnorm(10, 14.1, 1.6)
```

Less than 1% chance of observing a baby boy that was less than 10 pounds.

### Q68

```{r}
mu <- 185
sd <- 20
```

a. `r pnorm(195, mu, sd, lower = FALSE)`

b. `r pnorm(215, mu, sd) - pnorm(195, mu, sd)`

c. Probabilty of more than 215 for 1: `r (xx <- 1 - pnorm(215, mu, sd))` 

2 people independently `r xx^2`


## Section 3


### Q13

Using normal approximation:

```{r}
mu <- .2 * 25
sigma <- sqrt(.2 * .8 * 25)

pnorm(6.5, mu, sigma) - pnorm(3.5, mu, sigma)
```

Using Binom numbers

```{r}
pbinom(6, size = 25, prob = .2) - pbinom(3, size = 25, prob = .2)
```


### Q15

Normal approximation

```{r}
mu <- .4 * 20
sigma <- sqrt(.4 * .6 * 20)

1 - pnorm(9.5, mu, sigma)
```

Using Binom Numbers

```{r}
1 - pbinom(9, size = 20, prob = .4) 
```

### Q24


RH prob = .15

Blood bank get 92 donors

```{r}
mu <- .15*92
sigma <- sqrt(.15*.85*92)
```

a.
z score (10.5)

```{r}
(10.5 - mu)/sigma
```

probability using approximation

```{r}
pnorm(-.9635293)
```

exact using binomial

```{r}
pbinom(10, 92, .15)
```


b. 
zscore lower (14.5)

```{r}
(14.5 - mu)/sigma
```

score upper (20.5)
```{r}
(20.5 - mu)/sigma
```

answer:

```{r}

pnorm(1.956256) - pnorm(0.204385)
```


c. z-score (use 11.5 or fewer are rh NEGATIVE)

```{r}
(11.5 - mu)/sigma
```


```{r}

pnorm(-0.6715507)
```

### Q28

a.

exact

```{r}
1 - pbinom(20, 25, .6)
```

approximation

```{r}
mu <- .6 * 25
sigma <- sqrt(.6 * .4 * 25)

1 - pnorm(20.5, mu, sigma)
```


The approximation seems to slightly over estimate the real probability - very close though