---
title: "Homework Chapter 4"
format: html
---

```{r}
#| include: false
g <- glue::glue
library(ggplot2)
sapply(list.files(here::here('R'), full.names = T), source)
```

`r get_lecture_homework(4)`

## Section 4.1

### Q1

$S = \{1,2,3,4,5,6\} = \{E_1, E_2, E_3, E_4, E_5, E_6\}$

$A = \{2\}$

### Q2

$B = \{2,4,6\}$

### Q3

$C = \{3,4,5,6\}$

### Q4

$D = \{1,2,3,4\}$

### Q5

$E = \{2,4,6\}$

### Q6

$F = \{\}$

### Q28

5 balls - R1, R2, R3, Y1, Y2

These are the 20 events using a tree like diagram

```{r}
x <- c("R1", "R2", "R3" , "Y1", "Y2")



purrr::map(seq_along(x), function(.i) {
    out <- tibble::tibble(
        first_draw = x[.i],
        second_draw = x[-.i]
    )
}) |> 
purrr::list_rbind() |> 
flextable::flextable() |> 
flextable::merge_v() |> 
flextable::theme_box()
```

## Section 4.2

```{r}
get_lecture_homework(5)
```

### 4.2.7-4.2.10

7 Events, E7 has twice the probability.

Events 1-6 have probability *x*, E7 has probability 2x

$$
P(E_1) + P(E_2) + P(E_3) + P(E_4) + P(E_5) + P(E_6) + P(E_7) = 1
$$

$$
x + x + x +x +x +x + 2x = 1 = 8x
$$

$$
x = \frac{1}{8}
$$

#### 4.2.7

$$
x + x +x = \frac{3}{8}
$$

#### 4.2.8

$$
x  + x + x + 2x = \frac{5}{8}
$$

#### 4.2.9

$$
x + x = \frac{2}{8}
$$

#### 4.2.10

All three are not possible - therefore probability = 0

### 4.2.27

Part a:

Total must some to 1 therefore

$$
1 = .49 + .21 + .09 + ?
$$

Remaining probability = 0.21

Part b:

$$
.49 + .21  + .21 = .91 = 1 - .09
$$

### 4.2.36

Use tree diagram to show all simple events:

```{r}
x <- c("John", "Bill", "Ed", "Dave")

x_d <- expand.grid(first = x,
second = x,
third = x,
fourth = x) 

proper_combinations <- apply(x_d, MARGIN = 1, dplyr::n_distinct) == 4



all_comb <- x_d[proper_combinations, ] |> 
dplyr::mutate(probability = "1/24") |> 
dplyr::arrange(first, second, third, fourth) |> 
flextable::flextable() |> 
flextable::merge_v(j = -5)
all_comb
```

#### Part a

```{r}
all_comb |> 
flextable::bg(
    i = ~first=="Dave",
    j = ~probability,
    bg = "yellow"
)

```

6/24 = 1/4

#### Part b

```{r}
all_comb |> 
flextable::bg(
    i = ~first=="Dave" & second == "John",
    j = ~probability,
    bg = "yellow"
)

```

2/24

#### Part c

```{r}
all_comb |> 
flextable::bg(
    i = ~fourth=="Ed" ,
    j = ~probability,
    bg = "yellow"
)
```

6/24 = 1/4

## Section 4.3

```{r}
#| include: false
library(combinat)
```

### 4.3.1

$10 \times 8 = 80$

### 4.3.2

$4 \times 7 \times 3 = 84$

### 4.3.5-12

```{r}
perm <- function(n, k) {
    prod(tail(seq.int(n), k))
}

```

-   4.3.5: `r perm(5,3)`

-   4.3.6: `r as.integer(perm(10,9))`

-   4.3.7: `r perm(6,6)`

-   4.3.8: `r perm(20,1)`

-   4.3.9: `r choose(5,3)`

-   4.3.10: `r choose(10,9)`

-   4.3.11: `r choose(6,6)`

-   4.3.12: `r choose(20,1)`

### 4.3.13

$$
P^8_5 = 6720
$$

### 4.3.16

$$
4 \times 12 \times 4 = 192
$$

### 4.3.19

Part A: - 3 stages with n = 52 for each: - $52 \times 52 \times 52 = 140,608$

Part B:

-   3 stages with n = 52, 51, and 50, respectively
-   $52 \times 51 \times 50 = 132,600$

Part C: - 1st stage - pick any card = 52 - 2nd stage - 1 option same card as the previous person - 3rd stage - 1 option - same card as previous

-   Total options from part A
-   Probability of same card = $52/140608 = 0.00037$

Part D

-   Probability that all three students pick different cards:
-   Part B/Part A = $132600/140608 = 0.943$

### 4.3.21

-   Part A: order does not matter - $C^{52}_5 = \frac{52!}{(52 - 5)!5!} = 2,598,960$

-   Part B: Only 4 different combinations

-   Part C: 4/2598960 = 0.00000154

### 4.3.24

Part A

-   $C^5_2 = 30$

Part B:

-   1/30

### 4.3.24

### 4.3.26

-   Part A: $4 \times 3 \times 2 \times 1 = 24$

-   Part B: If Dave wins sprint there are $1 \times 3 \times 2 \times 1 = 6$ total ways he can win. Probability = 6/24

-   Part C: $1 \times 1 \times 2 \times 1 = 2$, Probability = 2/24

-   Part D: There are 3 options for first (any but Ed), 2 options for second (not and and not first), 1 option for 3rd left and Ed for fourth. Prob = 6/24


### 4.3.28

- Step 1 - how many ways can 5 problems go on the exam from 10 questions?
    - $C^{10}_5 = 252$

- Step 2 - howmany ways can you choose 5 problems from 6 questions?
   - $C^6_5 = 6$

Probability of all 5 being questions she knows is 6/252 = 0.024.

### 4.3.29

Total ways to draw the shapes: $12! = 479,001,600$

How many ways to group them when drawing them? $12 \times 2 \times 1 \times 9 \times 2 \times 1 \times 6 \times 2 \times 1 \times 3 \times 2 \times 1 = 31,104$

Probability = $31,104/479,001,600 = 0.00006494 = \frac{4!(3!)^4}{12!}$

## Section 4.4

### 4.4.1

-   a: $P(A\^c) = P(\\{E_2, E_4, E_5\\}) = .6$
-   b: $P(A \cap B) = P(\{E_1\}) = .2$
-   c: $P(B \cap C) = P(\{E_4\}) = .2$
-   d: $P(A \cup B) = P(\{E_1, E_2, E_3, E_4, E_5\}) = 1$
-   e: $P(B|C) = \frac{P(B\cap C)}{P(C)} = \frac{.2}{.4} = .5$
-   f: $P(A|B) = \frac{P(A\cap B)}{P(B)} = \frac{.2}{.8} = .25$
-   g: $P(A \cup B \cup C) = P(\{E_1, E_2, E_3, E_4, E_5\}) = 1$
-   h: $P(\left(A \cap B \right)^ c) = 1 - .2 = .8$

### 4.4.7 - 4.4.10

#### Q7

$$
P(A|B) = .1 = \frac{P(A\cap B)}{P(B)}
$$

$P(A\cap B) = .05$

#### Q8

Yes

#### Q9

No

#### Q10

No

### 4.4.14

-   a: Yes, $.3 \times .4 = .12$
-   b: 0 - they are mutually exclusive - use the addition rule
-   c: $P(A) = .3$
-   d: 0

### 4.4.27

$1 - 1/1000 \times 1/1000 = 1 - 1/1000000 = .999999$

### 4.4.29

-   a: $.7 \times .6 = .42$
-   b: Yes - probability of Mocha is the same regardless of where she goes
-   c: 30%
-   d: $P(A \cup B) = P(A) + P(B) - P(A\cap B) = .7 + .6 - .42 = .88$

### 4.4.32

$P(A) = .95$, $P(B) = .98$, $P(A\cap B) = .94$

-   a: $P(A \cup B) = P(A) + P(B) - P(A \cap B) = .99$
-   b: $1 - P(A \cup B) = .01$

### 4.4.35

-   a: $P(A) = 154/256 = 0.60$
-   b: $P(G) = 155/256 = 0.61$
-   c: $P(A \cap G) = 88/256 = 0.34$
-   d: $P(G|A) = 88/154 = 0.57$
-   e: $P(G|B) = 44/67 = 0.66$
-   f: $P(G|C) = 23/35 = 0.66$
-   g: $P(C|P) = 12/101 = 0.12$
-   h: $P(B^c) = 1 - 67/256 = 0.74$