---
title: "Homework Chapter 4"
format: html
---

```{r}
#| include: false
g <- glue::glue
library(ggplot2)
sapply(list.files(here::here('R'), full.names = T), source)
```

`r get_lecture_homework(4)`

## Section 4.1

### Q1

$S = \{1,2,3,4,5,6\} = \{E_1, E_2, E_3, E_4, E_5, E_6\}$

$A = \{2\}$

### Q2

$B = \{2,4,6\}$

### Q3

$C = \{3,4,5,6\}$

### Q4

$D = \{1,2,3,4\}$

### Q5

$E = \{2,4,6\}$

### Q6

$F = \{\}$

### Q28

5 balls - R1, R2, R3, Y1, Y2

These are the 20 events using a tree like diagram
```{r}
x <- c("R1", "R2", "R3" , "Y1", "Y2")



purrr::map(seq_along(x), function(.i) {
    out <- tibble::tibble(
        first_draw = x[.i],
        second_draw = x[-.i]
    )
}) |> 
purrr::list_rbind() |> 
flextable::flextable() |> 
flextable::merge_v() |> 
flextable::theme_box()
```

However, this assumes that R1 is different that R2 - which may not be the case. The tree is a lot simpler if we look at just the possibilities of two balls:

- R, R
- R, Y
- Y, Y
- Y, R


However, this is a little misleading because it makes it seem that all four possibilities are equally likely - which is not the case. We will revisit this in the future.
