---
title: "Homework Chapter 5"
format: html
---

```{r}
#| include: false
g <- glue::glue
library(ggplot2)
sapply(list.files(here::here('R'), full.names = T), source)
```

## Section 1




### Q12-Q16

$$
p(4) = 1 - (.1 + .3 + .4 + .1 + .05) = .05
$$



```{r}
d <- data.frame(
    x = 0:5,
    p_x = c(.1, .3, .4, .1, .05, .05)
)

ggplot(d, aes(x = x, y = p_x)) + 
    geom_bar(stat = "identity", width = 1)
```


```{r}
d <- d |> 
dplyr::mutate(
    `x * p_x` = round(x*p_x, 3),
    `x - mu` = round(x - sum(x*p_x), 3),
    `(x - mu)^2*p_x` = round(`x - mu`^2 * p_x, 3)
)

d |> flextable::flextable() |> flextable::autofit()

```

Mean:

```{r}
(d_mu <- sum(d$`x * p_x`))
```

Variance:

```{r}
(d_var <- sum(d$`(x - mu)^2*p_x`))
```


Standard Deviation:

```{r}
(d_std <- sqrt(d_var))
```


```{r}
ggplot(d, aes(x = x, y = p_x)) + 
    geom_bar(stat = "identity", width = 1) +
    geom_vline(xintercept = d_mu + c(-2,2)*d_std, linetype = "longdash")


```

This interval contains approximately 95% of the data since the data is relatively mound shaped. Interval:
```{r}
round(d_mu + c(-2,2)*d_std, 2)
```




### Q29


Probability of oil at any well $P(O) = .1$
Probability of no oil is therefore $P(O^c) = .9$
Probability that your first well has oil:

$$
p(1) = .1
$$

If you only hit on the second try then that means you missed on the first well and hit on the second. 

$$
p(2) = P(O^c) * P(O) = .09
$$

You can keep adding 'missed wells' with 3, etc.

$$
p(3) = P(O^c) * P(O^c) * P(O) = 0.081
$$


$$
p(n) = .1 \times .9^{n - 1}
$$


```{r}
plot(x = 1:5, y = .1 * .9^{1:5 - 1})
```


### Q34

```{r}
d_34 <- data.frame(
    x = 3:13,
    p_x= c(.03, .05, .07, .1, .14, .2, .18, .12, .07, .03, .01)

)
```

a. Mean
```{r}
(d_mean <- with(d_34, sum(x*p_x)))
```

b.

```{r}
(d_sd <- with(d_34, (x - d_mean)^2*p_x) |> sum() |> sqrt())
```


interval:

```{r}
(d_int <- round(d_mean + c(-2,2) * d_sd,2))
```

You can estimate it as about the data that runs between 4 and 11:
Approximately:
```{r}
sum(d_34$p_x[d_34$x >=4 & d_34$p_x <= 11])
```

Very similary to the expected 95%

## Section 2