---
title: "Midterm Review 02"
format: html
---

Note: some answers may disagree because of rounding in the 3rd or higher decimal point


```{r}
options(digits=3)
```

## Chapter 5

### Section 2

#### Q30

a.  `r pbinom(11, 20, .5)`
b.  `r pbinom(6, 15, .4)`
c.  `r 1 - pbinom(4, 10, .4)`
d.  `r 1 - pbinom(5, 15, .6)`
e.  `r pbinom(6, 10, .5) - pbinom(3, 10, .5)`

#### Q38

a.  This is binomial, n =100, p - .45
b.  The scores are not binomial - can take multiple values
c.  This is binomial, either you scored above average or not. p will be about 0.5 (assuming symmetric distribution) and n is about 45 (45% of 100)
d.  Not binomial, contiuous variable - maybe exponential?

#### Q44

a.  `r pbinom(3, 10, .5)`
b.  `r pbinom(6, 10, .5) - pbinom(3, 10, .5)`
c.  `r pbinom(6, 10, .5) - pbinom(3, 10, .5)`

#### Q46

a.  Odds to probability = 1/(2047 + 1) = 1/2048. Probability of 11 losses = $1/2^{11} = 1/2048$. Yes he is correct.

b.  $1/2^{13} = 1/8192$

## Chapter 6

### Section 1

#### Q5

0.5

#### Q6

0.15

#### Q7

0.5

#### Q8

0.45

#### Q17

a.  1/3
b.  1/3
c.  1/3

#### Q20

a.  `r pexp(5, 1/40)`
b.  `r 1 - pexp(50, 1/40)`
c.  `r pexp(30, 1/40)- pexp(20, 1/40)`
d.  `r pexp(120, 1/40)- pexp(-40, 1/40)`

### Section 2

#### Q49

```{r}
mu <- 69.5
sdev <- 3.5
```

```{r}
1 - pnorm(6*12, mu, sdev)
```

#### Q50

```{r}
pnorm(6*12 + 1, mu, sdev) - pnorm(5*12 +8, mu, sdev)
```

#### Q51

Probability someone is as tall as Trump or taller:

```{r}
1 - pnorm(6*12 + 2, mu, sdev)
```

#### Q52

Yes slightly unusual, as abou 43% of presidents are taller than 6 feet while only 24% of the population.

#### Q56

```{r}
pnorm(5, 10.5, 3)
```

#### Q59

a.  lower 25%

```{r}
qnorm(.25,  278, 12)
```

```         
upper 75%
```

```{r}
qnorm(.75,  278, 12)
```

b.  assume 6 months = 183 days

Probability is basically 0:

```{r}
pnorm(183,  278, 12)
```

#### Q61

```{r}
mu <- 1
sdev <- .15
```

a.  `r pnorm(1, mu, sdev)`
b.  `r pnorm(1.05, mu, sdev) - pnorm(.95, mu, sdev)`
c.  `r pnorm(.8, mu, sdev)`
d.  `r 1 - pnorm(1.45, mu, sdev)`. Yes unusual - probability is low.

### Section 3

#### Q17

```{r}
#.0446 in book 
1 - pnorm(59.5, .51*100, sqrt(100*.51*.49) )
```

#### Q19

```{r}
pnorm(155.5, .95*160, sqrt(160 * .95 * .05))
```

#### Q24

See [Homework](/homework/chapter-06.qmd#q24)
 
#### Q28

a. `r 1 - pbinom(20, 25, .6)`
b. `r 1 - pnorm(20.5, .6*25, sqrt(25*.6*.4))`

## Chapter 7

### Section 3

#### Q23

```{r}
mu <- 1110
sdev <- 80
se <- sdev/sqrt(400)
```

a. `r pnorm(1110, mu, se) - pnorm(1100, mu, se)`
b. `r 1 - pnorm(1120, mu, se)`
c. `r pnorm(900, mu, se)`

#### Q25

```{r}
mu <- 75878
sdev <- 4000

se <- sdev/sqrt(60)
```

a. Mean: 75878, and SE `r se`
b. `r glue::glue("[{round(mu - 2*se)}, {round(mu + 2*se)}]")`
c. `r 1 - pnorm(78000, mu, se)`
d. Yes this would be unusual, likely true average is not 75,878.


#### Q29

a. Yes appropriate because the population is normally distributed.
b. `r pnorm(20, 21, 2/sqrt(10))`
c. `r -qnorm(.001)*2/sqrt(10) +20`

#### Q32

See [Homework](/homework/chapter-07.qmd#q32)

### Section 5

#### Q26

Probability that $\hat p \ge .83$:

```{r}
1 - pnorm(.83, .8, sqrt(.8*.2/400))
```

$P(.76 \le \hat p \le .84)$:

```{r}
pnorm(.84, .8, sqrt(.8*.2/400)) - pnorm(.76, .8, sqrt(.8*.2/400))
```

#### Q27

```{r}
pnorm(.61, .5, sqrt(.5*.5/80)) - pnorm(.44, .5, sqrt(.5*.5/80))
```

#### Q28

```{r}
1 - pnorm(.8, .7, sqrt(.3*.7/50))
```


#### Q33
```{r}
p <- .13
q <- 1-p
n <- 55

se <- sqrt(p*q/55)
```

a. proportion: `r p`, standard error `r se`
b. `r pnorm(.2, p, se)`
c. `r 1 - pnorm(.35, p, se)`
d. [`r p - 1.96*se`, `r p + 1.96*se`] 


