---
title: "Midterm Review 01"
format: html
---

## Chapter 01

-   1.4.21, 1.4.22, 1.4.25

### Section 1.4

#### Q21 -Q 22

```{r}
d_1_4_21 <- c( 61, 93, 91, 86, 55, 63, 86, 82, 76, 57,
               94, 89, 67, 62, 72, 87, 68, 65, 75, 84)

```

```{r}
hist(d_1_4_21, 
    breaks = seq(52, length.out = 7, by = 8),  
    right = FALSE)

```

21. Has two peaks, one at about 65 and one at about 90
22. A bimodal shape is not usual. This could be because of two different classes that are differently prepared for the test.

#### Q25

```{r}
d_1_4_25 <- c(2.1, 4.4, 2.7, 32.3, 9.9, 9, 2, 6.6, 3.9, 1.6, 
14.7, 9.6, 16.7, 7.4, 8.2, 19.2, 6.9, 4.3, 3.3, 1.2,
4.1, 18.4, 0.2, 6.1, 13.5, 7.4, 0.2, 8.3, 0.3, 13,
14.1, 1, 2.4, 2.4, 18, 8.7, 24, 1.4, 8.2, 5.8,
1.6, 3.5, 11.4, 18, 26.7, 3.7, 12.6, 23.1, 5.6, 0.4)
```

Min, max:

```{r}
min(d_1_4_25)
max(d_1_4_25)
```

Range:

```{r}
range(d_1_4_25) |> diff()
```

Let's use 11 classes of width 3

```{r}
br <- seq(0, by = 3, length.out = 12)


x <- hist(d_1_4_25, breaks = br, right = FALSE)

tibble::tibble(
    interval_start = head(br, -1),
    interval_end = tail(br, -1),
    count = x$counts,
    frequency = round(x$density, 4)

) |> 
    knitr::kable(row.names =TRUE)
```

## Chapter 02

-   2.1.4, 2.1.9, 2.1.21
-   2.2.10
-   2.3.1, 2.3.6-14, 2.3.27

### Section 01

#### Q4

```{r}
d <- c(3,6,4,0,3,5,2)
```

Mean:

```{r}
mean(d)
```

Median:

```{r}
median(d)
```

Mode:

```{r}
table(d) |> sort() |> tail(1) |> names()
```

#### Q9

```{r}
d <- c(4.5, 3.2, 3.5, 3.9, 3.5, 3.9,
4.3, 4.8, 3.6, 3.3, 4.3, 4.2,
3.9, 3.7, 4.3, 4.4, 3.4, 4.2,
4.4, 4, 3.6, 3.5, 3.9, 4)
```

Mean:

```{r}
mean(d)
```

Median:

```{r}
median(d)
```

Dataset looks fairly symmetric, mean and median are very close to each other.

#### Q21

```{r}
d <- c(10.4, 10.7, 10, 9.2, 8.6, 8.9,
6.8, 10, 9.7, 9, 8.2, 
13, 6.8, 6.5, 6.4, 12, 
10.4, 14.3, 8.2, 15.6, 8.8)
```

Mean:

```{r}
mean(d)
```

Median:

```{r}
median(d)
```

Mode:

```{r}
table(d) |> sort() |> tail(4) |> names()
```

Data may be a little right skewed since the mean is to the right of the median.

Histogram (or dot plot) supports this statement

```{r}
hist(d)
```

### Section 2.2

#### Q10

```{r}
d <- c(1.28, 2.39 , 1.50 , 1.88 , 1.51)
```

Range:

```{r}
(d_range <- range(d) |> diff())
```

Sample Variance:

```{r}
var(d)
```

Sample Standard Deviation:

```{r}
(d_sd <- sd(d))
```

Mean:

```{r}
(xbar <- mean(d))
```

The various pieces of the sample variance formula:

```{r}
tibble::tibble(
    x = d,
    `x - xbar` = d - xbar,
    `(x - xbar)^2` = (d - xbar)^2

) |> knitr::kable()

```

Range/SD:

```{r}
d_range/d_sd
```

Range should be 4 x standard deviation, but it's about 2.5. Not a perfect approximation but not too suprising since there are only 5 data apoints.

### Section 2.3

#### Q1

```{r}
d <- c(5,2,3,6,1,2,4,5,1,3)
```

Range:

```{r}
(d_range <- range(d) |> diff())
```

Range approx of Std. Dev

```{r}
d_range/4
```

Std. Dev:

```{r}
(d_sd <- sd(d))
```

Pretty good approximation!

#### Q6 - Q11

```{r}
get_k <- function(interval, mean = 75, sd = 5) {
    abs((mean - interval)/sd)
}
```

6.  k = 1, total = 68%
7.  k = 2, total = 95%
8.  between 30 and 50 = 95/2 = 47.5%, between 50 and 60 = 68/2 = 34%, total = 47.5 + 34 = 81.5%
9.  \< 50 = 50%. 50 to 60 = 34% (see number 8), Greater than 60 = 100% - 34% - 50% = 16%.
10. Less than 60: 100% - 16% = 84%
11. 40 or more: 40 to 50 = 34%, greater than 50 = 50%, therefore 40 or more: 34 + 50 = 84%

#### Q12 - Q14

12. k = `r get_k(60)`, At least `r 1 - 1/get_k(60)^2`
13. k = `r get_k(65)`, At least `r 1 - 1/get_k(65)^2`
14. k = `r get_k(62.5)`, At least `r 1 - 1/get_k(62.5)^2`

#### Q27

```{r}
d <- c(56, 89, 51, 79, 58, 82, 52, 88, 52, 78, 69, 75, 77, 72, 71,
55, 87, 53, 85, 61, 93, 54, 76, 80, 81, 59, 86, 78, 71, 77)
```

Range:

```{r}
range(d) |> diff()
```

Range Approximatino of SD

```{r}
(range(d) |> diff())/4
```

Actual SD:

```{r}
sd(d)
```

Within 1 SD:

```{r}
perc_calc <- function(x, num_sds) {
    mean_x <- mean(x)
    sd_x <- sd(x)

    mean(x  < (mean_x + num_sds * sd_x) & 
        x  > (mean_x - num_sds * sd_x)
    )

}

perc_calc(d, 1)
```

Within 2 sds:

```{r}
perc_calc(d, 2)
```

Within 3 sds:

```{r}
perc_calc(d, 3)
```

The empirical rule isn't very accurate here - most likely to do the fact that the data is not mound shaped at all:

```{r}
hist(d)
```

## Chapter 04

-   4.1.13-15
-   4.2.11-15, 4.2.38, 4.2.40,
-   4.3.20, 4.3.25
-   4.4.16-18, 4.4.30, 4.4.36
-   4.5.7, 4.5.15

### Section 1

#### Q13-Q15

See notes from class [PDF of Notes](/homework/Midterm%2001%20Review.pdf)

### Section 2

#### Q11-Q15

E1 = E2 = .15, E3 = .4

E4 = 2E5

E1 + E2 + E3 + 2E5 + E5 = 1

.15 + .15 + .4 + 3E5 = 1

11. E5 = .1

    E4 = .2

12. .15 + .4 + .2 = .75

13. .15 + .4 = .55

14. P(E1, E2, E3, E4) = 0.9

15. 1 - .75 = .25

#### Q38 

1 \* 1/4 \* 1/4 = 1/16 = 4 \* (1/4 \* 1/4 \* 1/4)

#### Q40

a\. 114/220

b\. 25/220

c\. 156/220

### Section 3

#### Q20

2 \* 2 \* 2 = 8

#### Q25

a\. 6 \* 6 = 36

b\. 1/6 \* 1/6 = 1/36

c\. 5/6

### Section 4

#### Q16-17

-   Q16: a. 0, b. .8

-   Q17: a. .12/.4 = .3, b. no, c. Yes, P(B\|A) = P(B)

#### Q18

-   a\. 1

-   b\. 1

-   c\. 2/6

-   d\. 0

-   e\. 2/6

-   f\. 0

-   g\. 0

-   h\. 1

-   i\. 5/6

#### Q30

A - miss by first inspector

B - miss by second inspector

given: P(A) = .1, P(B\|A) = 5/10 = .5

Question is asking for A and B to be true

Use multiplication rule:

$P(A \cap B) = P(A)P(B|A) = .1 * .5 = .05$

#### Q36

a\. .71

b\. .29

c\. .35/.55

d\. .36/.45

e\. .35/.71

f\. .09/.29

### Section 5

#### Q7

$$
\begin{aligned}
P(A) &= P(A|B)P(A) + P(A|B^c)P(B^c) \\
 &= .6 * .3 + .5 * .4 \\
 &= 0.38
\end{aligned}
$$

#### Q15

a\.

$$
\begin{aligned}
P(D) &= .1 \\
P(D^c) &= .9 \\
P(N|D^c) &= .85/.9 = .94\\
P(N|D) &= .02/.1 = .2
\end{aligned}
$$

b\.

$$
\begin{aligned}
P(D|N) & = \frac{P(N|D) P(D)}{P(N)} \\
 &= \frac{P(N|D) P(D)}{P(N|D) P(D) + P(N|D^c) P(D^c)} \\
& = \frac{.2 *.1 }{.2 *.1  + .94* .9} \\
& = 0.023
\end{aligned}
$$

c\.

Same as B but

$$
P(D|N) = P(D \cap N)/P(N) = .02 / .87 = .023
$$

d\. $P(P|D^c)  = .05/.9 = 0.056$

e\. $P(N|D)  = .02/.1 = .2$

f\. The 20% false negative rate is a little high - it would miss 20% of disease.