---
title: "Homework Section 1.4"
format: html
---

## Question 17-20

```{r}
data <- c(1,2,1, 0, 2,
2,1,1,0,0,
2,2,1,1,0,
0,1,2,1,1)
```

```{r}
hist(data, breaks = -1:3, right = FALSE, freq = FALSE)
```

```{r}
table(data) |> tibble::as_tibble() |> 
gt::gt()
```

### 17

-   Greater than 1 (just 2s): 6
-   Total observations: `r length(data)`
-   Answer: `r 6/20`

### 18

-   1 - answer from 17
-   Answer: `r 1 - 6/20`

### 19

-   Answer: `r mean(data == 2)`

### 20

It doesn't look like there are any outliers - it is hard to detect them with just three different values. The distribution is relatively symmetric.

## Questions 21-22

```{r}
data <- c(61, 93, 91, 86, 55, 63, 86,82,76,57,94,89,67,62,72,87,68,65,75,84)
```

```{r}
hist_breaks <- seq(min(data), length.out = 6, by = 8)

hist(data, 
breaks = hist_breaks, right = FALSE)

```

I used these breaks `r hist_breaks`

### Q21

The shape of the scores seem pretty flat - with a small peak at \> 90.

### Q22

There seems to be an even distribution of student scores - meaning there aren't students who do really well or really poorly.

## Question 26

```{r}
data <- c(
    38,40,30,35,39,40,48,36,31,36,
    47,35,34,43,41,36,41,43,48,40,
    32,34,41,30,46,35,40,30,46,37,
    55,39,33,32,32,45,42,41,36,50,
    42,50,37,39,33,45,38,46,36,31
)
```

### a

```{r}
hist(data, 
   breaks = seq(30, length.out = 8, by = 5),
   right = FALSE)

```

### b

```{r}
mean(data >=35 & data < 45)
```

### c

```{r}
mean(data < 50)
```

## Question 33

```{r}
data <- c(.4, .6, .1, 1.1, 2.8,3.6,3.7,3.8,4,4,
4.4,4.5,5.3,5.3,5.5,5.8,5.9,6.1,6.4,6.6,
7,7.5,7.6,7.6,7.8,8.1,8.3,8.6,8.7,8.8,
8.8,8.9,8.9,9.1,9.2,9.4,9.6,9.7,9.8
)
```

### a

```{r}
hist(data, breaks = seq(0, by = 1, length.out = 11),
right = FALSE)

```

### b

Histogram is left skewed

### c

Maybe because as you move further away you get a chance to see more starbucks - there are only a few that are close to you but once you move out you can get a large swath of them.