---
title: "Lecture 11"
subtitle: Discrete Random Variables and Their Probability Distributions
format: html
date: 2023-10-05
---

```{r}
#| include: false
g <- glue::glue
library(ggplot2)
sapply(list.files(here::here('R'), full.names = T), source)
```

## Section 5.1 Discrete Random Variables and Their Probability Distributions

### Random Variables

For example, suppose you toss a die and measure *x* the number observed on the upper face. The variable *x* can take on any of six values - 1,2,3,4,5,6.

*x* is a random variable if the value that it assumes, corresponding to the outcome an experiment, is a chance or random event.

-   what are some examples of random variables?

-   This chapter focuses on *discrete* random variables.

### Probability Distribution

-   relative frequency distribution (sample)
-   **probability distribution** is the relative frequency distribution constructed for the entire population of measurements.
-   *p(x)* is the probability associated with value *x*

Requirements for a discrete probability distribution

-   $0 \le p(x) \le 1$

-   $\sum p(x) = 1$

### Example 5.1

Toss two fair coins and let x equal the number of heads observed. Find the probability distribution for x.

-   table with x, number of heads, simple events and p(x)

-   draw histogram with this information

### The Mean and Standard Deviation for a Discrete Random Variable

Population mean or *expected value*.

Suppose you the experiment is repeated a large number of times - say 4,000,000. How can we calculate the average?

Definition:

$$
\mu = E(x) = \sum x p(x)
$$

Repeat previous example but with 3 coin flips

You can also do the same thing for the population variance

$$
\sum(x - \mu)^2p(x)
$$

In both cases you sum over all possible values of *x*

### Example 5.2

A big box store is selling laptops, they only have 4 in stock. Previously, the marketing department had defined a probability distribution for the number of laptops bought as below. What is the expected number of laptops bought and the variance. How likely is it the 5 will be bought?

| x      | $p(x)$ | $xp(x)$ | $(x - \mu)^2$ | $(x - \mu)^2 p(x)$ |
|--------|--------|---------|---------------|--------------------|
| 0      | .1     |         |               |                    |
| 1      | .4     |         |               |                    |
| 2      | .2     |         |               |                    |
| 3      | .15    |         |               |                    |
| 4      | .1     |         |               |                    |
| 5      | .05    |         |               |                    |
| Totals |        |         |               |                    |

### Example 5.3

A lottery is conducted for a local charity, 8000 tickets are to be sold at \$10 each. The prize it a \$24,000 car. If you purchase two tickets, what is your expected gain?

What are the two possible gains? What is the probability of each?

What is the expected loss/gain?

```{r}

x <- 0:10
cost <- x*10

y <- -cost * (8000 - x)/8000 + (24000 - cost) * x/8000

plot(x, y)

```

### Example 5.4

An insurance company needs to know how much to charge for a \$100,000 policy insuring an event against cancellation due to inclement weather. The probability of inclement weather during the time of the event is assessed as 2 in 100. Once they find C, the cost of the policy to break event, they can add administrative costs and profit to this amount. Fin the value of C so that their expected gain is 0.

x = Gain, C = premium.

What are the two possibilities? What is the gain in each situation? What is the probability of each happening?

-   What has been an implicit assumption throughout (most) of these examples? is 2/100 for example, always true?

### Homework

```{r}
get_lecture_homework(12)
```