---
title: "Lecture 5"
subtitle: Events and the Sample Space
format: html
date: 2023-09-12
---

```{r}
#| include: false
g <- glue::glue
library(ggplot2)
sapply(list.files(here::here('R'), full.names = T), source)
```


## Section 4.2 Calculating Probabilties using Simple Events

The probabilit of an event _A_ is a measure of our believe that the event A will occur. 

$$
P(A) = \lim_{n\to\infty}
$$


 This is a limit of the relative frequency.

 - $P(A) = 0$
 - $P(A) = 1$


Each experiment results in one and only one single experiment.
- Probabiliy of a single event must like between 0 and 1
- The sum of the probabilities for all simple events in _S_ = 1.

If iwe can write down all simple events and find their individual prbabilities, we can find the probability of an event A:

- The **probability of an event A** is equal to the sume of the probabilies of the simple events contained in A

### Example 4.5

Toss two fair cons and record the outcome - find the probability of observing exactly one head.

What are the simple events?
What is the probability of observing exactly one head?


### Example 4.6

Proportinos of blood types A, B, AB and ) in the caucasian population in the US is .4, .11, .04, .45. What is the probability that a single caucasian person chosen randomly will have either type A or type AB Blood?


$$
P(\text{A or AB}) = P(\text{A}) + P(\text {AB}) = 0.4 + .04 = .44
$$


### Example 4.8

A candy dish contains one yellow and two red candies - close your eyes and choose two candies - what is the probability that both are red?

- Need to list all simple events via tree:

```{r}
x <- c("R1", "R2", "Y1")

expand.grid(first_draw = x,
second_draw = x) |>  
dplyr::arrange(first_draw) |> 
flextable::flextable() |> 
flextable::merge_v()
```



### Homework

```{r}
get_lecture_homework(5)
```

Answers: [Chapter 4 - Section 2](/homework/chapter-04.qmd#section-4.2)