---
title: "Lecture 8/9"
subtitle: Rules for Calculating Probabilities
format: html
date: 2023-09-21
---

```{r}
#| include: false
g <- glue::glue
library(ggplot2)
sapply(list.files(here::here('R'), full.names = T), source)
```

## Section 4.4 Rules for Calculating Probabilities

Let *A* and *B* be two events defined in the sample space *S*

-   union - $A \cup B$ - is the event that either *A* or *B* or both occur

-   intersection - $A \cap B$ - both *A* and *B* occur

-   complement - $A^c$ is the event that *A* does not occur

```{r}

blank_venn <- function(title = "") {
    A <- 1:5
    B <- 3:10

    ggvenn::ggvenn(
        list(A, B), c("A", "B"), 
    show_percentage = FALSE, fill_alpha = 0, show_elements = TRUE) +
    labs(title = paste0("  ", title)) +
    theme(plot.background = element_rect(color = "black", linewidth = 1))
}


blank_venn("A union B")
blank_venn("A intesect B")
blank_venn("A complement")
```

### Example 4.17

Two fair cons are tossed and the outcome is recorded.

-   *A*: Observe at least one head
-   *B*: Observe at least one tail

Define A, B, $A \cup B$, $A \cap B$, $A^c$

-   What are the events and what are the probabilities?

-   These concepts can be extended to more than two events (e.g. $A \cup B \cup C$)

### Calculating probabilities for Unions and Complements

-   The addition rule:

$$
P (A \cup B) = P(A) + P(B) - P(A \cap B)
$$

```{r}
blank_venn("Addition Rule")
```

-   mutually exclusive: $P(A \cap B) = 0$

-   Rule for complements: $P(A^c) = 1 - P(A)$

### Example 4.18

-   Drilling two exploratory wells:

-   *A* neither well produces oil or gas - 80%

-   *B* exactly one well produces oil or gas - 18%

-   *C* Both wells produce oil or gas - 2%

-   Find $P(A \cup B)$ and $P(B \cup C)$

### Calculating Probabilities for Intersections

-   Two events *A* and *B* are independent if and only if the probability of event *B* is not influenced or changed by the occurrence of event *A* or vice versa.

-   Color blindness/gender vs tossing a dice

-   General Multiplcation Rule

$$
P(A \cap B) = P(A)P(B|A) = P(B)P(A|B)
$$

$P(A|B)$ is read as probability of A given B - assuming B occurs. If the two events are independent then $P(A|B) = P(A|B^c) = P(A)$

### Example 4.20

In a color preference experiment, eight toys are placed in a container. The toys are identical except fo color - two are red, and six are green. A child is asked to choose two toys at random. What is the probability that the child chooses the two red toys?

-   Use a tree diagram fo first and second choice
-   $A = \text{(R on first choice) } \cap \text{ (R on second choice)}$

### Conditional Probabilities

The conditional probability of event *A* given that event *B* has occured is:

$$
P(A|B) = \frac{P(A\cap B)}{P(B)} \text{  if  } P(B) \neq 0
$$

$$
P(B|A) = \frac{P(A\cap B)}{P(A)} \text{  if  } P(A) \neq 0
$$

### Colorblindness, continued

|                        | Men($B$) | Women($B^c$) | Total |
|------------------------|----------|--------------|-------|
| Colorblind ($A$)       | .04      | .002         | .042  |
| Not Colorblind ($A^c$) | .47      | .488         | .958  |
| Total                  | .51      | .49          | 1     |

-   If you pick a random person what is the chance they are colorblind?

-   If you pick a random male, what is the chance he is colorblind?

-   What is the chance a woman is?

### The multiplication rule for independent events

If two events *A* and *B* are independent, the probability that *both* A and B occur is:

$$
P(A \cap B) = P(A)P(B)
$$

Same thing for more than two events:

$$
P(A \cap B \cap C) = P(A)P(B)P(C)
$$

### Coin tosses  In a Row

-   What the probability that the first toss is a heads, second toss is a tails, third toss is a heads?

-   What about all three tosses being heads?

### Checking for independence

Two events are said to be independent if and only if either:

$$
P(A\cap B) = P(A)P(B)
$$

or

$$
P(B|A) = P(B)
$$

### Example 4.22

Respondents were asked about the cost of a college education . The respondents were classified according to whether they currently had a chaild in college and whther they thought the loan burden for most college students is too high, the right amount or too little.

|                         | Too High (A) | Right Amount (B) | Too Little (C) |
|-------------------------|--------------|------------------|----------------|
| Child in College (D)    | .35          | .08              | .01            |
| No Child in College (E) | .25          | .20              | .11            |

Is A independent of D?

Also Remember the question about shooting free throw - is making the second one independent of making the first one?

|            | Hit Second | Miss Second |
|------------|------------|-------------|
| Hit First  | .49        | .21         |
| Miss First | .21        | .09         |

### Homework

```{r}
get_lecture_homework(9)
```

Answers: [Chapter 4 - Section 4](/homework/chapter-04.qmd#section-4.4)