---
title: "Lecture 8/9"
subtitle: Rules for Calculating Probabilities
format: html
date: 2023-09-21
---

```{r}
#| include: false
g <- glue::glue
library(ggplot2)
sapply(list.files(here::here('R'), full.names = T), source)
```

## Section 4.4 Rules for Calculating Probabilities

Let *A* and *B* be two events defined in the sample space *S*

-   union - $A \cup B$ - is the event that either *A* or *B* or both occur

- intersection - $A \cap B$ - both *A* and *B* occur

- complement - $A^c$ is the event that *A* does not occur


```{r}

blank_venn <- function(title = "") {
    A <- 1:5
    B <- 3:10

    ggvenn::ggvenn(
        list(A, B), c("A", "B"), 
    show_percentage = FALSE, fill_alpha = 0, show_elements = TRUE) +
    labs(title = paste0("  ", title)) +
    theme(plot.background = element_rect(color = "black", linewidth = 1))
}


blank_venn("A union B")
blank_venn("A intesect B")
blank_venn("A complement")
```

### Example 4.17

Two fair cons are tossed and the outcome is recorded. 

- *A*: Observe at least one head
- *B*: Observe at least one tail

Define A, B, $A \cup B$, $A \cap B$, $A^c$

- What are the events and what are the probabilities?


- These concepts can be extended to more than two events (e.g. $A \cup B \cup C$)


### Calculating probabilities for Unions and Complements

- The addition rule:

$$
P (A \cup B) = P(A) + P(B) - P(A \cap B)
$$

```{r}
blank_venn("Addition Rule")
```

- mutually exclusive: $P(A \cap B) = 0$


- Rule for complements: $P(A^c) = 1 - P(A)$

### Example 4.18

- Drilling two exploratory wells:

- *A* neither well produces oil or gas - 80%
- *B* exactly one well produces oil or gas - 18%
- *C* Both wells produce oil or gas - 2%

- Find $P(A \cup B)$ and $P(B \cup C)$

### Calculating Probabilities for Intersections

- Two events *A* and *B* are independent if and only if the probability of event *B* is not influenced or changed by the occurrence of event *A* or vice versa.


- Color blindness/gender vs tossing a dice

- General Multiplcation Rule

$$
P(A \cap B) = P(A)P(B|A) = P(B)P(A|B)
$$


$P(A|B)$ is read as probability of A given B - assuming B occurs. If the two events are independent then $P(A|B) = P(A|B^c) = P(A)$

### Example 4.20

in a color preference experiment, eight toys are placed in a container. The toys are identical except fo color - two are red, and six are green. A child is asked to choose two toys at random. What is the probability that the child chooses the two red toys?

- Use a tree diagram fo first and second choice
- $A = \text{(R on first choice) } \cap \text{ (R on second choice)}$



### Homework

```{r}
get_lecture_homework(9)
```

Answers: [Chapter 4 - Section 4](/homework/chapter-04.qmd#section-4.4)