---
title: "Lecture 4"
subtitle: Events and the Sample Space
format: html
date: 2023-09-05
---

```{r}
#| include: false
g <- glue::glue
library(ggplot2)
sapply(list.files(here::here('R'), full.names = T), source)
```

## Introduction

- Basic concepts of probability so that we can draw conclusions about a sampled population

- Example: Toss a single coin, probability of a heads (or tails) is 50%

-  If you toss 10 times and see 10 heads. If you have a fair coin, the probability that you'd see 10 heads is very low. It's likely the coin is biased

- Two ways 
    -  probability of even when the population is known, 
    - probabiltiy of inference given you observe a sample from the population

## 4.1 Events and the sample space

- An **experiment** is the process by which an observation or measurement is obtained.

Examples:

- Recording a test grade
- Measuring daily rainfall
- Recording a persons's opinion
- Any others??

When an experiment is performed we observe an outcome called a **simple event** - Often written as $E_i$

Example -

What are the simple events that can happen when a 6 sided die is rolled?

- $E_1$ = 1
- $E_2$ = 2
- ...
- $E_6$ = 6


An **event** is a collectino of simple events. for example:

- _A_: Observe an odd number: $A = \{E_1, E_3, E_5\}$
- _B_: Observe a number less than 4: $B = \{E_1, E_2, E_3\}$


Two events are **mutually exclusive** if when one event occurs the otherr cannot - and vise versa.

Note that $A$ and $B$ are not mutuall exclussive - why?

- What are some example of mutually exlcussive events

- The set of all simple events i called the sample space , _S_


Venn Diagram: (see page 129 of book)
- Venn Diagram of die tossing, including all simple event and A/B


Some examples:

### Ex 4.2

Toss a simple coin - what are the simple events that define the sample space?

### Ex. 4.3 & 4.4


Blood Types (4 Simple Events)

These are mutually exclusive

- what about adding Rh factor - (positive/negative)
- Use a tree diagram to show how to come up with the simple events


```{r}
get_lecture_homework(4)
```

Answers: [Chapter 4 - Section 1](/homework/chapter-04.qmd#section-4.1)
