---
title: Lecture 19
subtitle: The Sampling Distribution of the Sample Proportion

format: html
date: 2023-11-14
---

```{r}
#| include: false
g <- glue::glue
library(ggplot2)
sapply(list.files(here::here('R'), full.names = T), source)
```

## Section 7.5 the Sampling Distribution of the Sample Proportion

Can you estimate the true population proportion *p* from a sample?

-   Binomial distribution has mean of what?
-   What about standard deviation?

### Properties of the Sampling Dsitribution of the Sample Proportion of $\hat p$

-   If a random sample of *n* observations is selected from a binomial population with parameter *p*, then the sampling distribution of the sample proportion

$$
\hat p = \frac x n
$$

will have a mean

$$
p
$$

and a standard deviation

$$
\text{SE}(\hat p) = \sqrt{ \frac {pq}{n}}
$$

When sample size o*n* is large, the sampling distribution of $\hat p$ can be approximated by a normal distribution. The approximation will be adequate if $np > 5$ and $nq > 5$.

### Example 7.10

-   500 parents asked about the importance of sports for boys and girls. Of the parents interviewed, 60% agreed that boys and girls should have equal opportunities to participate in sports. Describe the sampling distribution of the sample proportion $\hat p$ of parents who agree that boys and girls should have equal opportunities.

$$ 
\text{SE}(\hat p) = \sqrt{ \frac {pq}{n}} \approx \sqrt{ \frac {\hat p \hat q}{n}} = \sqrt{ \frac {.6 * .4}{500}} = .022
$$

Create 95% Interval

-   95% of the time, $\hat p$ will fall within 2SE $\approx .044$ of the unknown value of $p$

-   This is the first time we have *estimated* a parameter from the data - and estimated the uncertainty around the estimate.

-   We have generally reversed the situation and assumed a population proportion $p$ and calculated the probability of how likely or unlikely a sample proportion $\hat p$ might be.

### Example 7.11

Suppose someone claims that the proportion $p$ of parents in the population is actually .55. What is the probability of observing a sample proportion as large as or larger than $\hat p = .6$

$$ 
\text{SE}(\hat p) = \sqrt{ \frac {pq}{n}}  = \sqrt{ \frac {.55 * .45}{500}} = .0222
$$

```{r}
xplot <- seq(.45, .65, length.out = 100)


plot(xplot, dnorm(xplot, mean = .55, sd = .0222), type = "l")

```

$$
z = \frac{\hat p - p}{\sqrt{pq/n}} = \frac{.6 - .55}{.0222} = 2.25
$$

$$
P(\hat p > .6) \approx P(z > 2.25) = 1 - .9878 = .0122
$$

With continuity correction:

$$
z_1 = \frac{(.6 - .5/500) - .55}{.0222} = 2.20
$$

### Homework

```{r}
get_lecture_homework(22)
```

Answers: [Section 7.5](/homework/chapter-07.qmd#section-5)