---
title: "Lecture 15/16"
subtitle: Probability Distributions for Continous Random Variables
format: html
date: 2023-11-02
---

```{r}
#| include: false
g <- glue::glue
library(ggplot2)
sapply(list.files(here::here('R'), full.names = T), source)
```

## Section 6.3 The normal Approximation to the Binomial Probability Distribution

-   The binomial formala produces lengthy calculations and the tables are available only for certain values of n and p.
-   You can use the normal approximation to the binomial probability distribution

### The normal approximation

-   Let *x* be a binomial random variable with n trials and probability p of success. The probability distribution of x is approximated using a normal curve with:

-   $\mu = np$ and $\sigma = \sqrt{npq}$.

-   This approximation is adequate as long as *n* is large and *p* is not too close to 0 or 1.

### Figure 6.17

-   n = 25, p = 0.5
-   $\mu = ?$
-   $\sigma = ?$
-   compare the areas under the curve to the areas in the bar
-   Example the curve between 7.5 and 10.5 is approximately $P(8 \le x \le 10)$

```{r}
tibble::tibble(
    x = 0:25,
    y = dbinom(x, size = 25, prob = .5) 

) |> 
ggplot(aes(x = x, y = y)) +
geom_bar(stat = "identity", fill = "white", color = "black") +
stat_function(fun = function(x) dnorm(x, mean = .5*25, sd = sqrt(.5*.5*25)), color = "blue")  +
scale_x_continuous(breaks = 0:25) +
labs(x = "x", y = "p(x)", caption = "blue line is normal distribution")
```

Use Binomial Probability? $$
P(8 \le x \le 10) = .190
$$

Use Normal approximation (subtract .5 from lower end and add .5 to upper end)

$$
P(7.5 \le x \le 10.5) = .1891
$$

-   z-score

The 1/2 half adjustment is called a continuity correction and helps account for the effect that you are approximating a discrete random variable with a continuous one.

Rule of thumb:

-   The normal approximation to the binomial probabilities will be adequate if both:
-   $np >5$ and $nq >5$

### Figure 6.18

-   n = 25, p = .1
-   $\mu = ?$
-   $\sigma = ?$

```{r}
tibble::tibble(
    x = -3:25,
    y = dbinom(x, size = 25, prob = .1) 

) |> 
ggplot(aes(x = x, y = y)) +
geom_bar(stat = "identity", fill = "white", color = "black") +
stat_function(fun = function(x) dnorm(x, mean = .1*25, sd = sqrt(.1*.9*25)), color = "blue")  +
scale_x_continuous(breaks = -3:25) +
geom_vline(xintercept = 0) +
labs(x = "x", y = "p(x)", caption = "blue line is normal distribution")
```

### Example 6.12

The reliability of an electrical fuse it he probability that a fuse, chosen at random from pdouction will function under its designed conditions. A random sample of 1000 fuses was tested and x = 27 defectives were observed. Calculate the approximate probablity of observing 27 or more defectives, assuming that the fuse reliability is .98.

-   $\mu = np = 1000*.02 = 20$

-   $\sigma = \sqrt{npq} = 1000 * .02 * .98 = 19.6$

-   Check if normal approximation is appropriate?

-   Create z score for value of 26.5

-   Normal approximation yields .0708

Using exact:

```{r}
1 - pbinom(27, size = 1000, prob = .02, lower.tail = TRUE) |> round(4)
```

### Example 6.12

A soda manufacturer was fairly certain that her brand had a 10% share of the market. In a survey involving 2500 soda drinkers, x = 211 preferred her brand. If the 10% figure is correct, find the probability of observing 211 or fewer consumers who perfer her brand of soda.

-   $\mu = 250$

-   $\sigma = 15$

-   Z score for 211.5 = -2.57

$$
P(x \le 211) \approx P( z \lt -2.57) = .0051
$$

Exact binomial probability:

```{r}
pbinom(211, size = 2500, prob = .1) |> round(4)
```

What happened??

\
\
\
\
\
\
\
\
\
\
\
\
\

Either observed an unusual sample even though p = 0.1 OR the sample reflects the value of p is less than .1 and perhaps closer to 21/2500 = .08

### Homework

```{r}
get_lecture_homework(19)
```

Answers: [Section 6.3](/homework/chapter-06.qmd#section-3)