---
title: "Lecture 10"
subtitle: Bayes' Rule
format: html
date: 2023-10-03
---

```{r}
#| include: false
g <- glue::glue
library(ggplot2)
sapply(list.files(here::here('R'), full.names = T), source)
```

## Section 4.5 Bayes' Rule

-   Remember the experiment on color blindness:

|                        | Men($B$) | Women($B^c$) | Total |
|------------------------|----------|--------------|-------|
| Colorblind ($A$)       | .04      | .002         | .042  |
| Not Colorblind ($A^c$) | .47      | .488         | .958  |
| Total                  | .51      | .49          | 1     |

: Rates of color blindeness in men and women

-   Event $B$: person is a man
-   Event $B^c$: person is a woman\

Taken together it's the full sample space and are mutually exclusive.

-   What if we want to describe all color blind people? We have two types, those in $B$ and those in $B^c$

-   How do we represent color blind people in $B$?

    -   $A \cap B$

-   How do we present color blind people in $B^c$?

    -   $A \cap B^c$

-   How do we represent all color blind people:

-   $(A \cap B) \cup (A \cap B^c)$\

Using the addition rule:

$$
P(A) = P(A \cap B) + P(A \cap B^c)
$$

Why is:

$$
P((A \cap B) \cap (A \cap B^c)) = 0
$$

Let's now assume there are more than 2 *mutually exclusive and exhaustive* categories - we can extend the rule using *k* populations $S_1, S_2, S_3, \dots, S_k$.

$$
A = (A \cap S_1) \cup (A \cap S_2) \cup (A \cap S_3) \cup \dots \cup (A \cap S_k)
$$

Then $$
P(A) = P(A \cap S_1) + P(A \cap S_2) + P(A \cap S_3) + \dots + P(A \cap S_k)
$$

-   Show venn diagram that is on page 159

Remember also that:

$$
P(A \cap S_i) = P(S_i)P(A|S_i)
$$

### The Law of Total Probability

$$
P(A) = P(S_1)P(A|S_1) + P(S_2)P(A|S_2) + P(S_3)P(A|S_3) + \dots + P(S_k)P(A|S_k) 
$$

### Bayes' Rule

One useful use of this is when you need to find the conditional probability that A occurred given B $P(A|B)$, but you know $P(B|A)$. This is common with screening tests.

-   Event $D$ is that you have the disease
-   Event $T$ is that the test is positive

A test is often characterized by the true positive rate $P(T|D)$, true negative rate $P(T^c|D^c)$, false positive rate $P(T|D^c)$ and the false negative rate $P(T^c|D)$. You can see all these values depend on a state of disease. This is because when they are creating a new test, they can find people with the disease and without the disease - i.e. they are known to have the disease.

-   However, what we really care about is the probability you have the disease given a positive test $P(D|T)$.

Let's imagine a scenario where you have a test that is pretty good when it detects $P(T|D) = .99$ and doesn't have a ton of false positives $P(T|D^c) = .01$. Does this seem like a good test? \
\
\
The answer is that it depends. Let's see why:

Recall that:

$$
P(D)(P(T|D)) = P(D\cap T) = P(T)P(D|T)
$$

Then:

$$
P(D|T) = \frac{P(T|D)P(D)}{P(T)}
$$

Where

$$
P(T) = P(D)P(T|D) + P(D^c)P(T|D^c)
$$

Plugging in for $P(T)$:

$$
P(D|T) = \frac{P(T|D)P(D)}{P(D)P(T|D) + P(D^c)P(T|D^c)}
$$

What can we fill in?

$P(T|D) = .99$ and $P(T|D^c) = .01$

$$
P(D|T) = \frac{.99P(D)}{.99P(D) + .01P(D^c)}
$$

Let's say we are trying to detect the common cold. What if it's the winter? There is a lot of cold out there, assuming you are feeling sick then there is a 40% chance it's the cold.

Therefore: $P(D) = .4$ and $P(D^c) = 1 - .4 = .6$ So what's the probability that you have the cold given there is a positive test?

$$
P(D|T) = \frac{.99 \times .4}{.99 \times .4 + .01 \times .6} = 0.985 = 98.5\%
$$

However, what if it's the summer - there isn't a lot of cold going around out there. Not a lot of people have the cold so $P(D) = .05$ and $P(D^c) = .95$.

$$
P(D|T) = \frac{.99 \times .05}{.99 \times .05 + .01 \times .95} = 0.839 = 83.9\%
$$

The posterior probability $P(D|T)$ changes even though the test information is the same. It changes because the prior ($P(D)$) matters.\
\
\
\
\
\
This can be generalized to:

$$
P(S_i|A) = \frac{P(S_i)P(A|S_i)}{\sum_{j=1}^kP(S_j)P(A|S_j)}
$$

for $i = 1, 2, \dots, k$

### Homework

```{r}
get_lecture_homework(11)
```

Answers: [Chapter 4 - Section 5](/homework/chapter-04.qmd#section-4.5)

## Midterm Review

-   1.4.21, 1.4.22, 1.4.25
-   2.1.4, 2.1.9, 2.1.21
-   2.2.10
-   2.3.1, 2.3.6-14, 2.3.27
-   4.1.13-15
-   4.2.11-15, 4.2.38, 4.2.40,
-   4.3.20, 4.3.25
-   4.4.16-18, 4.4.30, 4.4.36
-   4.5.7, 4.5.15