---
title: "Lecture 1"
subtitle: Welcome & Relative Frequency Histograms
format: html
date: 2023-08-29
---


::: {.content-hidden when-format="html"}

## Welcome 
- background
- questions: 
    - communication via email? Chat?
- [syllabus.qmd] 
:::


## 1.4 Relative Frequency Histograms



```{r}
#| label: tbl-1-9
#| tbl-cap: Birthweights of 30 Full-Term Newborn Babies

tbl_1_9 <- c(7.2, 7.8, 6.8, 6.2, 8.2, 8, 8.2, 5.6, 8.6, 7.1, 8.2, 7.7, 
    7.5, 7.2, 7.7, 5.8, 6.8, 6.8, 8.5, 7.5, 6.1, 7.9, 9.4, 9, 7.8, 
    8.5, 9, 7.7, 6.7, 7.7) |> 
    matrix(ncol = 5,
    byrow = TRUE)

knitr::kable(tbl_1_9,  col.names = NA)
```

- Definitions
    - class
    - class boundaries
    - class width
    - class frequency
    - relative frequency histogram


### Choosing Number of Classes

```{r}
#| label: tbl-1-10
#| tbl-cap: Choosing the number of classes 
tibble::tibble(
    `25` = 6, 
    `50` = 7,
    `100` = 8,
    `200` = 9,
    `500` = 10
) |> 
knitr::kable()
```
    

### Choosing class width

Range of Data

$$
9.4 - 5.6 = 3.8 \\
$$

Divide and round

$$
3.8/8 = .475 \approx .5
$$


```{r}
data_freq <- cut(tbl_1_9, seq(5.6, by = .5, length.out = 9), right = FALSE) |> 
    table()  |> 
    as.data.frame() |> 
    tibble::as_tibble()

colnames(data_freq)[1] <- "Class Boundaries"

data_freq[,1] |> 
    knitr::kable(align = "c")

```

- Method of left inclusion

### Tally Data in each of Boundaries

```{r}
data_freq |> 
    transform(
        `Relative Frequency` = glue::glue("{Freq}/30")
    ) |> 
    knitr::kable(align = 'r')
```

### Draw Histogram 

```{r}
library(ggplot2)
ggplot(mapping = aes(as.vector(tbl_1_9))) + 
    geom_histogram(
            breaks = seq(5.6, by = .5, length.out = 9),     closed = "left",
            color = "white") +
    labs(x = "Birth Weights", 
         y = "Frequency")
```

```{r}

data_freq |> 
   transform(breaks = seq(5.6, by = .5, length.out = 8),
                 `Relative Frequency` = Freq/sum(Freq))   |> 
ggplot(mapping = aes(x= breaks, y = Relative.Frequency)) + 
    geom_bar(stat = "identity",
            color = "white", just = 0, width = .5) +
            labs(x = "Birth Weights", 
         y = "Relative Frequency")
```

### Example 1.11

```{r}
#| label: tbl-1-12
#| tbl-cap: Number of Visits in Typical Week for 25 Customers

tbl_1_12<- c(6,7,1,5,6,4,6,4,6,8,6,5,6,3,4,5,5,5,7,6,3,5,7,5,5) |> 
    matrix(ncol = 5,
    byrow = TRUE)

knitr::kable(tbl_1_12,  col.names = NA)
```


- Use intgers as class definitions

```{r}
data_freq <- cut(tbl_1_12, seq(0.5, by = 1, length.out = 9), right = FALSE) |> 
    table()  |> 
    as.data.frame() |> 
    tibble::as_tibble()

colnames(data_freq)[1] <- "Class Boundaries"


data_freq |> 
    transform(
        `Relative Frequency` = Freq/length(tbl_1_12)
    ) |> 
    knitr::kable(align = 'r')
```



```{r}

ggplot(mapping = aes(as.vector(tbl_1_12))) + 
    geom_histogram(
            aes(y = after_stat(density)),
            breaks = seq(0.5, by = 1, length.out = 9),     closed = "left",
            color = "white") +
            labs(x = "Visits", 
         y = "Relative Frequency")

```


### Using Histograms

- data location
- data shape (skewed left, skewed right)
- proportion/probability


::: {.content-hidden when-format="html"}
- *proportion*: how many of the observed are higher
- *probability*: if you pick a random one what would you get?
:::


### Making histograms in Excel

Download a sample Exel file [here](lecture-01-histograms.xlsx)


### Looking forward

- sample vs population
- if you look at 30 more birth weights will you have same histogram?


## Other Resources

- [I'm not a real statistican and you can be one too](https://statsepi.substack.com/p/im-not-a-real-statistician-and-you)


::: {.content-hidden when-format="html"}
## Stretch
- Histograms from different data sources showing bimodal, left and right skew
- ask class to guess what data could have generated this data.
:::


## Homework

```{r}
get_lecture_homework <- function(lecture_number) {
    schedule_raw <- readxl::read_xlsx(here::here('resources','semester-planning.xlsx'))


    subset(schedule_raw, `Lecture Number` == lecture_number, select = exercises, drop = TRUE)

}

```


`r get_lecture_homework(1)`

Answers: [Section 1.4](/homework/section-1_4.qmd)
